---
title: "In-Hospital Cardiac Arrest & ICU Outcomes Audit"
author: "Hussein Kassir"
date: "`r Sys.Date()`"
output:
  pdf_document:
    toc: true
  html_document:
    toc: true
    toc_float: true
    theme: flatly
    highlight: tango
    number_sections: true
subtitle: "Modelled on ICNARC National Cardiac Arrest Audit — MIMIC-IV Demo Dataset"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)

library(tidyverse)
library(lubridate)
library(broom)
library(knitr)

# Load all outputs
cohort_final       <- read_csv("outputs/cohort_final.csv")
discrepancy_log    <- read_csv("outputs/discrepancy_log.csv")
care_unit_breakdown <- read_csv("outputs/care_unit_breakdown.csv")
age_breakdown      <- read_csv("outputs/age_breakdown.csv")
admission_breakdown <- read_csv("outputs/admission_breakdown.csv")
model_results      <- read_csv("outputs/logistic_regression_results.csv")
model_data         <- read_csv("outputs/model_data.csv")
```

# Executive Summary

This report presents findings from an audit of in-hospital cardiac arrest 
and ICU outcomes, modelled on the methodology of ICNARC's National Cardiac 
Arrest Audit (NCAA) and Case Mix Programme (CMP). The audit was conducted 
using the MIMIC-IV Clinical Database Demo (v2.2), an openly available 
dataset of 100 de-identified critical care patients from Beth Israel 
Deaconess Medical Center, Boston.

**Key findings:**

- One cardiac arrest case was identified in the demo dataset (ICD-10: I462),
  admitted as an emergency to the Coronary Care Unit, with in-hospital death 
  confirmed.
- Across the full ICU cohort of 100 patients (140 stays), overall in-hospital 
  mortality was **13%**.
- The Coronary Care Unit recorded the highest mortality rate at **38.5%**, 
  followed by the Neuro Surgical ICU at **33.3%**.
- ICU length of stay was the only statistically significant predictor of 
  in-hospital death in the risk model (OR 1.30, 95% CI 1.12–1.58, p=0.002).
- 51 data quality discrepancies were identified and logged, of which 9 were 
  rated High severity and 42 Medium severity.

> **Note:** This audit was developed as a portfolio project to demonstrate 
> clinical audit data management and analytical skills. The MIMIC-IV Demo 
> dataset contains 100 patients and results are not statistically powered 
> for clinical inference. The pipeline is designed to scale to the full 
> MIMIC-IV dataset.


# Background & Methods

## Audit Scope

This audit was designed to mirror the structure and objectives of ICNARC's 
National Cardiac Arrest Audit (NCAA), which monitors the incidence, care, 
and outcomes of in-hospital cardiac arrest across NHS hospitals in England, 
Wales, and Northern Ireland. A secondary analysis of broader ICU outcomes 
was conducted across the full cohort, modelled on ICNARC's Case Mix 
Programme (CMP).

## Dataset

The **MIMIC-IV Clinical Database Demo (v2.2)** was used as the data source. 
MIMIC-IV is a large, freely available database of de-identified health data 
from patients admitted to critical care units at Beth Israel Deaconess 
Medical Center between 2008 and 2019. The demo version contains 100 patients 
and is openly available via PhysioNet without credentialing requirements.

## Cohort Definition

**Cardiac arrest cohort:**

Patients were included if they had a recorded diagnosis of cardiac arrest 
using the following codes:

- ICD-10: I46, I460, I461, I462, I469
- ICD-9: 4275

Patients were excluded if aged under 18 or flagged with a high-severity 
data quality discrepancy.

**Full ICU cohort:**

All patients with at least one ICU stay in the dataset were included in the 
secondary analysis, subject to the same age and QC exclusion criteria.

## Outcome Measures

The primary outcome was **in-hospital mortality**, defined as 
`hospital_expire_flag = 1` in the admissions table. Secondary outcomes 
included ICU and hospital length of stay, care unit distribution, and 
admission type breakdown.

## Analytical Pipeline

All data processing, quality control, cohort definition, and analysis were 
conducted in R (v4.x) using a reproducible, version-controlled pipeline 
maintained on GitHub. Data quality checks were modelled on ICNARC's 
discrepancy resolution process, with all flagged records logged with a 
severity rating and resolution status.

# Data Quality Summary

## Overview

Data quality checks were conducted across all four core tables: `patients`, 
`admissions`, `diagnoses_icd`, and `icustays`. Three categories of checks 
were applied: completeness, plausibility, and consistency. All flagged 
records were logged in a structured discrepancy log with a severity rating 
and resolution status, replicating the query-raising process used in live 
clinical audits.

## Completeness

```{r completeness-table}
tibble(
  Table = c("patients", "admissions", "admissions", "admissions", "admissions"),
  Variable = c("dod", "deathtime", "edregtime / edouttime", 
               "discharge_location", "marital_status"),
  `% Missing` = c("69%", "94.5%", "33.8%", "15.3%", "4.4%"),
  Interpretation = c(
    "Expected — only populated for deaths",
    "Expected — only populated for in-hospital deaths",
    "Expected — not all patients admitted via emergency department",
    "Flagged — unknown discharge destination, audit-relevant",
    "Minor gap, low audit relevance"
  )
) %>% kable()
```

## Plausibility

No implausible values were identified for age, ICU length of stay, or 
admission/discharge dates. All 100 patients were adults (≥18), all ICU 
lengths of stay were positive, and no discharge dates preceded admission dates.

## Consistency

Four records showed death time recorded after discharge time by a few hours. 
Investigation revealed this reflects a known MIMIC data convention where 
discharge time is recorded as midnight (00:00:00) and death time is logged 
precisely later that evening. These were documented as known artefacts and 
excluded from the discrepancy log.

Nine records were flagged for ICU stay timestamps falling outside the 
hospital admission window by more than six hours, likely reflecting 
date-shift artefacts in the MIMIC de-identification process.

## Discrepancy Log Summary

```{r discrepancy-summary}
discrepancy_log %>%
  group_by(check_type, severity) %>%
  summarise(n_records = n(), .groups = "drop") %>%
  arrange(severity, check_type) %>%
  kable(col.names = c("Check Type", "Severity", "Records Flagged"))
```

In a live audit setting, High severity discrepancies would be queried back 
to the submitting unit for clarification or correction before inclusion in 
analysis. Medium severity records would be flagged for review and included 
with a data quality caveat.


# Cardiac Arrest Cohort

## Cohort identification

```{r ca-cohort-flowchart}
tibble(
  Step = c(
    "All MIMIC-IV Demo patients",
    "Cardiac arrest ICD code identified",
    "Excluded — under 18",
    "Excluded — high severity QC flag",
    "Final analysis cohort"
  ),
  N = c(100, 1, 0, 0, 1)
) %>% kable(col.names = c("Step", "N"))
```

## Patient Profile

```{r ca-patient}
cohort_final %>%
  select(
    gender, anchor_age, admission_type, admission_location,
    first_careunit, los, hospital_los, in_hospital_death, icd_code
  ) %>%
  mutate(
    across(everything(), as.character),
    in_hospital_death = ifelse(in_hospital_death == "TRUE", "Yes", "No")
  ) %>%
  rename(
    Gender                = gender,
    Age                   = anchor_age,
    `Admission Type`      = admission_type,
    `Admission Location`  = admission_location,
    `Care Unit`           = first_careunit,
    `ICU LOS (days)`      = los,
    `Hospital LOS (days)` = hospital_los,
    `In-Hospital Death`   = in_hospital_death,
    `ICD Code`            = icd_code
  ) %>%
  pivot_longer(everything(), names_to = "Variable", values_to = "Value") %>%
  kable()
```

## Findings & Limitations

The single cardiac arrest case identified was a female patient aged 89, 
admitted as an emergency via the emergency department and managed in the 
Coronary Care Unit. In-hospital death was confirmed, with an ICU length 
of stay of 4.9 days and hospital length of stay of 4.8 days. The arrest 
was coded as ICD-10 I462 (cardiac arrest due to underlying cardiac condition).

> **Limitation:** The MIMIC-IV Demo dataset contains only 100 patients, 
> yielding a single cardiac arrest case. Outcome statistics for this cohort 
> are not meaningful at this scale. The identification, validation, and 
> reporting pipeline is fully operational and designed to scale to the full 
> MIMIC-IV dataset, which contains over 300,000 hospital admissions.


# Full ICU Cohort Analysis

## Overview

A secondary analysis was conducted across all 100 patients with at least 
one ICU stay in the dataset, yielding 140 ICU stays. This analysis mirrors 
the Case Mix Programme's approach to benchmarking critical care outcomes 
across participating units.

## In-Hospital Mortality

```{r mortality}
tibble(
  Metric = c("Total patients", "In-hospital deaths", 
             "Survivors", "Mortality rate"),
  Value  = c("100", "13", "87", "13%")
) %>% kable()
```

## ICU Length of Stay

```{r los}
tibble(
  Metric = c("Mean ICU LOS (days)", "Median ICU LOS (days)",
             "Minimum ICU LOS (days)", "Maximum ICU LOS (days)"),
  Value  = c("3.68", "2.16", "0.02", "20.5")
) %>% kable()
```

The median ICU length of stay was 2.16 days, with considerable variation 
(range 0.02–20.5 days), reflecting the heterogeneity of case mix typical 
of a mixed critical care population.

## Breakdown by Care Unit

```{r care-unit}
care_unit_breakdown %>%
  rename(
    `Care Unit`       = first_careunit,
    `ICU Stays`       = n_stays,
    `Deaths`          = n_deaths,
    `Mortality (%)`   = mortality_pct,
    `Mean LOS (days)` = mean_los
  ) %>%
  kable()
```

The Coronary Care Unit recorded the highest mortality rate at 38.5%, 
followed by the Neuro Surgical ICU at 33.3%. The Cardiac Vascular ICU 
recorded zero deaths across 25 stays, likely reflecting a higher proportion 
of elective surgical admissions with lower baseline risk.

## Breakdown by Age Band

```{r age-band}
age_breakdown %>%
  rename(
    `Age Band`      = age_band,
    `Patients`      = n_patients,
    `Deaths`        = n_deaths,
    `Mortality (%)` = mortality_pct
  ) %>%
  kable()
```

Mortality was highest in the 60–79 age band at 21.4%. The apparent lower 
mortality in the 80 and over group (12.5%) may reflect survivor bias — 
only the fittest older patients are admitted to ICU — and should be 
interpreted with caution given the small sample size.

## Breakdown by Admission Type

```{r admission-type}
admission_breakdown %>%
  rename(
    `Admission Type` = admission_type,
    `Patients`       = n_patients,
    `Deaths`         = n_deaths,
    `Mortality (%)`  = mortality_pct
  ) %>%
  kable()
```

Elective and surgical same-day admissions recorded zero mortality, 
consistent with their lower acuity. Direct emergency admissions had the 
highest mortality at 50%, though this is based on only two patients and 
should not be interpreted as a reliable estimate.

# Risk-Adjusted Outcomes

## Logistic Regression Model

A logistic regression model was fitted to identify predictors of in-hospital 
mortality across the full ICU cohort. Predictor variables were selected to 
reflect those used in ICNARC's risk prediction models: age, sex, admission 
type (emergency vs non-emergency), and ICU length of stay.

```{r logistic-regression}
model_results %>%
  filter(term != "(Intercept)") %>%
  mutate(
    term = recode(term,
      "age"       = "Age (per year)",
      "emergency" = "Emergency admission",
      "female"    = "Female sex",
      "los"       = "ICU length of stay (per day)"
    ),
    `Significant` = ifelse(p.value < 0.05, "Yes", "No")
  ) %>%
  rename(
    `Predictor`   = term,
    `Odds Ratio`  = estimate,
    `Std. Error`  = std.error,
    `Z Statistic` = statistic,
    `P Value`     = p.value,
    `95% CI Low`  = conf.low,
    `95% CI High` = conf.high
  ) %>%
  kable()
```

## Key Findings

- **ICU length of stay** was the only statistically significant predictor 
  of in-hospital death (OR 1.30, 95% CI 1.12–1.58, p=0.002). Each 
  additional day in ICU was associated with a 30% increase in the odds 
  of in-hospital death, likely reflecting greater illness severity in 
  patients with prolonged stays.
- **Age** showed a positive but non-significant association with mortality 
  (OR 1.04, p=0.11), consistent with the expected direction but 
  underpowered at this sample size.
- **Emergency admission** was associated with higher odds of death 
  (OR 2.91) but with very wide confidence intervals (0.61–22.5), 
  reflecting insufficient statistical power.
- **Female sex** showed a non-significant protective association 
  (OR 0.36, p=0.198).

## Observed vs Expected Mortality

```{r oe-ratio}
tibble(
  Metric = c("Observed deaths", "Expected deaths", "O/E ratio"),
  Value  = c("13", "13.0", "1.00")
) %>% kable()
```

The observed-to-expected (O/E) ratio of 1.00 indicates perfect calibration 
of the model on this dataset. In a live audit context, an O/E ratio 
substantially above 1.0 would indicate higher-than-expected mortality and 
trigger a quality improvement review, while a ratio below 1.0 would suggest 
better-than-expected performance. This mirrors the Risk-Adjusted Mortality 
Indicator (RAMI) used in ICNARC's Case Mix Programme.

> **Note:** The O/E ratio here reflects in-sample model fit rather than 
> external validation. In a scaled audit, the model would be trained on a 
> derivation dataset and validated on a separate cohort.


# Limitations & Recommendations

## Limitations

**Sample size:** The MIMIC-IV Demo dataset contains 100 patients, yielding 
a single cardiac arrest case. All statistics derived from this cohort should 
be interpreted as illustrative of the pipeline methodology rather than 
clinically meaningful findings. The full MIMIC-IV dataset contains over 
300,000 hospital admissions and would provide adequate statistical power 
for a fully powered audit.

**Single-centre data:** MIMIC-IV draws exclusively from Beth Israel 
Deaconess Medical Center in Boston. A national audit such as ICNARC's NCAA 
draws from hundreds of participating units across England, Wales, and 
Northern Ireland, enabling meaningful inter-unit benchmarking. Single-centre 
data cannot replicate this.

**ICD coding limitations:** Cardiac arrest identification relied solely on 
ICD diagnostic codes. In clinical practice, coding completeness varies 
across units and cardiac arrest may be undercoded, particularly when it 
occurs as a complication rather than a primary diagnosis. A live audit would 
supplement ICD coding with direct data entry by clinical staff.

**Risk model variables:** The logistic regression model included only four 
predictor variables. ICNARC's own risk prediction models incorporate a much 
richer set of physiological variables collected at ICU admission, including 
acute physiology scores, ventilation status, and primary diagnosis category. 
These variables were not included here due to the complexity of extracting 
and cleaning chartevents data within the scope of this project.

**Date shifting:** MIMIC-IV applies date shifting to all timestamps as part 
of its de-identification process. This introduced apparent inconsistencies 
in some timestamp comparisons, particularly between ICU stay times and 
hospital admission windows. These were identified through QC checks and 
documented in the discrepancy log.

## Recommendations

1. **Scale to full MIMIC-IV dataset** — applying this pipeline to the full 
   dataset would yield a statistically powered cardiac arrest cohort and 
   enable meaningful benchmarking analysis.

2. **Enrich the risk model** — incorporating physiological variables from 
   `chartevents` (e.g. GCS, heart rate, blood pressure at ICU admission) 
   would produce a more clinically credible risk adjustment model, closer 
   to ICNARC's RAMI methodology.

3. **Automate discrepancy reporting** — in a live audit setting, the 
   discrepancy log could be extended to generate automated query letters 
   to participating units, flagging specific records and requesting 
   clarification or correction.

4. **Extend to NCAA variables** — where derivable from MIMIC-IV, 
   additional NCAA variables such as arrest rhythm (shockable vs 
   non-shockable) and ROSC status could be extracted from clinical notes 
   or procedure codes to more closely replicate the full NCAA dataset 
   specification.
   
